---
title: "final_sdgs"
format: html
editor: visual
---

## Overview

For my final project for the SDG analytics in R, I chose to focus on the relationship between SDG 9 and 10. Recently for a capstone essay in my healthcare analytics course, I briefly studied the relationship between stronger infrastructure and equality in both less developed countries (LDCs) and more developed countries (MDCs).

Within this project I seek to understand the relationship between goal 9 and 10 using data visualization, logistic regression, and the SDR 2023 dataset. I seek to recognize the countries with strong and weak SDG 9 progress, and correlate the two goals, to understand my question, "What is the relationship between SDG 9 and SDG 10 in a global context?" and my secondary question, "How does this relationship differ in LDCs and MDCs?

This research will consist of 4 main sections, first, understanding my data , cleaning it, and basic visualization to set a basis for my analysis. Secondly, I will create 3-4 data visualizations that give insights to goal 9 and 10 progress in different countries . Third, I will create a logistic regression model to determine correlation between goal 9 and 10 in different countries, and finally I will conclude with my analysis of the results and closing thoughts, limitations and future works.

## Section 1: Libraries and exploratory vis

```{r}
#| echo: false
#| #libraries & loading in data
library(tidyverse)
library(here)
library(janitor)
library(plotly)
library(rnaturalearth)
library(leaflet)
library(sf)
library(vembedr)
library(ggpubr)
library(ggcorrplot)

sdr_data <- read_csv(here("data/SDR-2023-Data.csv"))
```

```{r}
## data cleaning + basic exploration
sdr_data <- sdr_data %>% 
  clean_names()

table(sdr_data$goal_7_dash, sdr_data$regions_used_for_the_sdr)

```

```{r}
#exploration
sdr_data %>% 
  select(goal_9_score, goal_10_score)
```

## Section 2: Data Visualization

```{r}
###First vis
ggplot(data = sdr_data, aes(x= goal_9_score, y = goal_10_score, color = regions_used_for_the_sdr)) +
  geom_point()
```

From just this basic visualization there seems to be no correlation between goal 9 and 10, and that's fine! I'm not looking for breakthrough correlation, simply looking at what the data says about my answer. However, it's important to continue fully exploring, because one scatterplot is a good introduction, but cannot tell the full story. However, what this scatterplot can tell me is that Sub-Saharan Africa regions have a lower goal 9 score, and a moderate range in their goal 10 score. OECD regions have both a high goal 9 and goal 10 score. This doesn't appear to be a huge correlation though, it just tells us what countries are doing better or worst at goal 9 and 10, however there could be a correlation between regions with a high count of MDCs or LDCs, which tells me it's worth investigating. So, for my next visualization I would have to define regions that are MDC or LDC.

To define if a region is an MDC or LDC in the cleanest, easiest way, I could use another goal as a cue. SDG 1, no poverty, is a good contender, and is used by the UN (in part with goal 3, 4, 8, and others) to distinguish MDC's and LDCs. Although not a perfect picture of what separates the country classifications, for the purpose of my project and the MAIN goal being to see if there is a correlation between sdg 9 and 10, I am willing to accept this limitation, and will establish this further in my conclusion.

```{r}
ggplot(data = sdr_data, aes(x= goal_9_score, y = goal_1_score, color = regions_used_for_the_sdr)) +
  geom_point()
```

```{r}
ggplot(data = sdr_data, aes(x= goal_10_score, y = goal_1_score, color = regions_used_for_the_sdr)) +
  geom_point()
```
