---
title: "final_sdgs"
format: html
editor: visual
---

## Overview

For my final project for the SDG analytics in R, I chose to focus on the relationship between SDG 9 and 10. Recently for a capstone essay in my healthcare analytics course, I briefly studied the relationship between stronger infrastructure and equality in both less developed countries (LDCs) and more developed countries (MDCs).

Within this project I seek to understand the relationship between goal 9 and 10 using data visualization, logistic regression, and the SDR 2023 dataset. I seek to recognize the countries with strong and weak SDG 9 progress, and correlate the two goals, to understand my question, "What is the relationship between SDG 9 and SDG 10 in a global context?" and my secondary question, "How does this relationship differ in LDCs and MDCs?

This research will consist of 4 main sections, first, understanding my data , cleaning it, and basic visualization to set a basis for my analysis. Secondly, I will create 3-4 data visualizations that give insights to goal 9 and 10 progress in different countries . Third, I will create a logistic regression model to determine correlation between goal 9 and 10 in different countries, and finally I will conclude with my analysis of the results and closing thoughts, limitations and future works.

## Section 1: Libraries and exploratory vis

```{r}
#| echo: false
#| #libraries & loading in data
library(tidyverse)
library(here)
library(janitor)
library(plotly)
library(rnaturalearth)
library(leaflet)
library(sf)
library(vembedr)
library(ggpubr)
library(ggcorrplot)

sdr_data <- read_csv(here("data/SDR-2023-Data.csv"))
```

```{r}
## data cleaning + basic exploration
sdr_data <- sdr_data %>% 
  clean_names()

table(sdr_data$goal_7_dash, sdr_data$regions_used_for_the_sdr)

```

```{r}
#exploration
sdr_data %>% 
  select(goal_9_score, goal_10_score)
```

## Section 2: Data Visualization

```{r}

```
