---
title: "final_sdgs"
format: html
editor: visual
---

## Overview

For my final project for the SDG analytics in R, I chose to focus on the relationship between SDG 9 and 10. Recently for a capstone essay in my healthcare analytics course, I briefly studied the relationship between stronger infrastructure and equality in both less developed countries (LDCs) and more developed countries (MDCs).

Within this project I seek to understand the relationship between goal 9 and 10 using data visualization, linear regression, and the SDR 2023 dataset. I seek to recognize the countries with strong and weak SDG 9 progress, and correlate the two goals, to understand my question, "What is the relationship between SDG 9 and SDG 10 in a global context?" and my secondary question, "How does this relationship differ in LDCs and MDCs?

This research will consist of 4 main sections, first, understanding my data , cleaning it, and basic visualization to set a basis for my analysis. Secondly, I will create 3-4 data visualizations that give insights to goal 9 and 10 progress in different countries . Third, I will create a linear regression model to determine correlation between goal 9 and 10 in different countries, and finally I will conclude with my analysis of the results and closing thoughts, limitations and future works.

## Libraries and exploratory vis

```{r}
#| echo: false
#| #libraries 
#| 
```

```{r}

```
