---
title: "week2.2"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(vembedr)
```

```{r}
sdr_data <- read_csv(here("data/SDR-2023-Data.csv"))
```

```{r}
sdr_data <- sdr_data %>% 
  clean_names()

```

```{r}
ggplot(data = sdr_data, aes(x= goal_3_score, y = goal_4_score)) +
  geom_point()
```

```{r}
ggplot(data = sdr_data, aes(x= goal_3_score, y = goal_4_score, color = regions_used_for_the_sdr)) +
  geom_point()
```

```{r}
ggplot(data = sdr_data, aes(x= goal_3_score, y = goal_4_score, color = goal_5_score)) +
  geom_point()
```

```{r}
##Challenge 1
ggplot(data = sdr_data, aes(x= goal_3_score, y = goal_6_score, color = goal_5_score)) +
  geom_point()
```

Goal 3 and 6 have a strong positive correlation. Goal 5 also has a strong positive correlation, however less so. (Goal 5 gender equality, goal 6 sustainable access to drinking water, goal 3 good health and well-being.)

```{r}
ggplot(sdr_data, aes(x= goal_3_score, y = country)) +
  geom_bar(stat = "identity")
```

```{r}
lac_sdr_data <- sdr_data %>% 
  filter(regions_used_for_the_sdr == "LAC")
```

```{r}
ggplot(lac_sdr_data, aes(x= goal_3_score, y = country)) +
  geom_bar(stat = "identity") 

```

```{r}
ggplot(sdr_data, aes(x= goal_3_score, y = country)) +
  geom_bar(stat = "identity") +
  facet_wrap(~regions_used_for_the_sdr, scales = "free_y") + 
  theme(axis.text.y = element_text(size = 4))
```

```{r}
##challenge 2 bar chart
c<- sdr_data %>% 
  filter(regions_used_for_the_sdr == "MENA")

ggplot(mena_sdr_data, aes(x= goal_5_score, y = country)) +
  geom_bar(stat = "identity") 
```

```{r}
ggplot(sdr_data, aes(x = goal_4_score)) +
  geom_histogram()
```

```{r}
ggplot(sdr_data, aes(x = goal_4_score, fill=regions_used_for_the_sdr)) + 
  geom_histogram(bins=33, color="black")
```

```{r}
##challenge 3

ggplot(sdr_data, aes(x = goal_5_score, fill=regions_used_for_the_sdr)) + 
  geom_histogram(bins=33, color="black")
```

Countries in the OECD (European countries, parts of America, etc) had higher SDG score for SDG 5 (gender quality.) There seems to be no / limited correlation between count (population size) and score. Sub-Saharan and MENA regions peform the lowest in SDG 5.
